An Integral Projection Modeling Approach to Demographic Consequences of Multiple Partner Mutualisms

### 2022-11-17 Manuscript/manuscript_draft_ALI.tex
Manuscript updates -- Ali

Mutualisms are species interactions in which all involved participants benefit. They are among the most widespread species interactions\cite{Chamberlain2014,BoucherDouglasH.1985}, and are notoriously context dependent\cite{Bronstein1994,Chamberlain2014,Frederickson2013}, often deteriorating into other types of interactions, including parasitism \cite{Rodriguez-Rodriguez2017,Song2020,Mandyam2014,Thrall2007}.
\ali{The benefits from these interactions are called rewards, but they require investment of resources into the interaction}{I'm honestly not sure if I need to explain this idea? Do I need to include more basic mutualism background? I have tried to shift this so that I cover a small amount and then jump into the idea of partner diversity in mutualisms instead}. 
While mutualism is defined at the level of a species pair (+/+) these interactions are embedded in multi-species communities, such that a focal species (the focal mutualist) may engage with multiple partner species.
%Understanding why these relationships deteriorate or persist is important as mutualism is a cornerstone in maintaining the biodiversity of many systems\cite{Isbell2011}.

Partner diversity in mutualisms often leads to variable fitness benefits for the focal mutualist\cite{Afkhami2014, Palmer2010}, due to differences in partner quality\cite{Bascompte2019,Stanton2013,Frederickson2013,Jones2015}, differences in the types of benefits offered\cite{Kiers2003,Afkhami2014}, and direct interactions between partners\cite{Sun2019,Heath2009,Heath2014,Grutter2003}.
%Interactions between partners, partner niche overlap, temporal resource variation, and differences in partner quality, functional benefit, and competitive ability all impact the fitness benefits to the focal mutualist\cite{stanton2013, Afkhami2014,Boucher1982}.
For these reasons, pairwise studies cannot be accurately used to predict the outcomes of multi-species mutualisms\cite{Palmer2010, Stanton2013. Chamberlain2014, Song2020}.
Studies on the effects of diverse multi-species mutualism systems are necessary to help us understand the demographic effects of partner diversity in mutualisms.

Multi-species mutualism partner diversity causes varied demographic effects on the population of the focal mutualist which can be explained by a number of mechanisms.
Partners can vary in many ways: in some cases, the quality varies leading to some true mutualist partners, some freeloaders, and even some parasites\cite{Bronstein1994, Afkhami2014,Song2020,West2007,Frederickson2013}; in other cases, the function of the partner can vary, each offering a different type of reward to the focal mutualist \cite{Stanton2003}.
When there is a consistent partner hierarchy based on benefits offered, a more diverse sample of the partner community may be more likely to include the highest quality one\cite{Frederickson2013} (the one which offers the focal mutualist the most benefit).
When the fitness of a focal mutualist with multiple partners is equal to the fitness of a focal mutualist with only the highest quality partner, as shown in Figure ***, sampling effect can explain the positive effects of partner diversity\cite{Batstone2018}. 
Partners can also vary functionally, meaning each partner offers a different type of reward to the focal mutualist\cite{Stachowicz2005,Bronstein2006,Stanton2003}.
When this leads to a higher fitness for the focal mutualist than any single interaction, complementarity explains the benefits of partner diversity\cite{Batstone2018}. 
Partners can also have asynchronous responses to the environment, either spatially\cite{Ollerton2006} or temporally\cite{Alarcon2008}.
Multiple partners can act as a 'portfolio' leading to more even fitness benefits across environmental or temporal heterogeneity, as shown in Figure ***, through the portfolio effect\cite{Batstone2018,Lazaro2022}.

When multiple partners are present, partner turnover is also important to consider, as partner identity has been shown to impact benefits in mutualisms\cite{Djieto-Lordon2005}. 
Turnover can happen at different timescales\cite{Oliveira1999,Horvitz1986} from minutes to years. 
The degree and order of partner turnover can have important impacts on the fitness of the focal mutualist.
High degrees of partner turnover, also known as oportunism, may offer buffering in the face of environmental variation\cite{Hegland2009}, while low degrees of partner turnover, also known as partner fidelity, can increase the benefits received in a mutualism\cite{Sachs2004}.
The direction of turnover can also have a significant impact, particularly if the most beneficial partner changes across the ontogeny of the focal mutualist's life\cite{Fonseca2003}.
For example, susceptibility to enemies can change across life stages\cite{Boege2005}, so can defense traits, indicating that the focal mutualist benefits the most when defensive partners align with focal mutualist development\cite{Djieto-Lordon2005}.

Defensive ant-plant mutualisms are common interactions in which plants provide extra-floral nectar and/or ``housing'' to ants which in turn defend them from herbivores\cite{Bronstein1998}. 
Ants are highly ubiquitous\cite{Schultheiss2022} and vary significantly in abundances\cite{Schultheiss2022}, types of rewards offered\cite{Beattie1985}, and quality of interaction with their plant partners\cite{Rico-Gray1989,Palmer2010}, making these interactions great systems to study the effects of partner diversity within mutualisms. 
While these interactions have been well studied\cite{Ness2006,Beattie1985,Schultheiss2022} in the literature, few have considered how diversity within ant defender guilds\cite{Stanton2013} or temporal fluctuations in partner interactions\cite{} affect the overall benefits of mutualism for the plant partner.

>>>>

Mutualisms are species interactions in which all involved participants benefit. They are among the most widespread species interactions\cite{Chamberlain2014,BoucherDouglasH.1985}, and are notoriously context dependent\cite{Bronstein1994,Chamberlain2014,Frederickson2013}, often deteriorating into other types of interactions, including parasitism \cite{Rodriguez-Rodriguez2017,Song2020,Mandyam2014,Thrall2007, Bahia2022}.
The benefits from these interactions are called rewards, but they require investment of resources into the interaction\cite{Bronstein2001}.
The presence of mutualists can shift the available habitat, fitness, or resource uptake of a partner\cite{Stachowicz2001,West2007}.
While mutualism is defined at the level of a species pair (+/+) these interactions are embedded in multi-species communities, such that a focal species (the focal mutualist) may engage with multiple partner species\cite{Stanton2013, Boucher1982}.

Partner diversity in mutualisms often leads to variable fitness benefits for the focal mutualist\cite{Afkhami2014, Palmer2010}, due to differences in partner quality\cite{Bascompte2019,Stanton2013,Frederickson2013,Jones2015, Ness2006}, differences in the types of benefits offered\cite{Kiers2003,Afkhami2014}, and direct interactions between partners\cite{Sun2019,Heath2009,Heath2014,Grutter2003}.
For these reasons, pairwise studies cannot be accurately used to predict the outcomes of multi-species mutualisms\cite{Palmer2010, Stanton2013. Chamberlain2014, Song2020}.
Studies on the effects of diverse multi-species mutualism systems are necessary to help us understand the demographic effects of partner diversity in mutualisms\cite{Bascompte2019}.

Multi-species mutualism partner diversity causes varied demographic effects on the population of the focal mutualist which can be explained by a number of mechanisms.
Partners can vary in many ways: in some cases, the quality varies leading to some true mutualist partners, some freeloaders, and even some parasites\cite{Bronstein1994,Bronstein2001a,Afkhami2014,Song2020,West2007,Frederickson2013,Jones2015}; in other cases, the function of the partner can vary, each offering a different type of reward to the focal mutualist \cite{Stanton2003}.
When there is a consistent partner hierarchy based on benefits offered, a more diverse sample of the partner community may be more likely to include the highest quality one\cite{Frederickson2013} (the one which offers the focal mutualist the most benefit).
When the fitness of a focal mutualist with multiple partners is equal to the fitness of a focal mutualist with only the highest quality partner, as shown in Figure ***, sampling effect can explain the positive effects of partner diversity\cite{Batstone2018}. 
Partners can also vary functionally, meaning each partner offers a different type of reward to the focal mutualist\cite{Stachowicz2005,Bronstein2006,Stanton2003}.
When this leads to a higher fitness for the focal mutualist than any single interaction, complementarity explains the benefits of partner diversity\cite{Batstone2018}. 
Partners can also have asynchronous responses to the environment, either spatially\cite{Ollerton2006} or temporally\cite{Alarcon2008}.
Multiple partners can act as a 'portfolio' leading to more even fitness benefits across environmental or temporal heterogeneity, as shown in Figure ***, through the portfolio effect\cite{Batstone2018,Lazaro2022}.

When multiple partners are present, partner turnover is also important to consider, as partner identity has been shown to impact benefits in mutualisms\cite{Djieto-Lordon2005}. 
Turnover can happen at different timescales\cite{Oliveira1999,Horvitz1986} from minutes to years. 
The degree and order of partner turnover can have important impacts on the fitness of the focal mutualist.
High degrees of partner turnover, also known as oportunism, may offer buffering in the face of environmental variation\cite{Hegland2009}, while low degrees of partner turnover, also known as partner fidelity, can increase the benefits received in a mutualism\cite{Sachs2004}.
The direction of turnover can also have a significant impact, particularly if the most beneficial partner changes across the ontogeny of the focal mutualist's life\cite{Fonseca2003}.
For example, susceptibility to enemies can change across life stages\cite{Boege2005,Barton2010}, so can defense traits, indicating that the focal mutualist benefits the most when defensive partners align with focal mutualist development\cite{Djieto-Lordon2005}.

Defensive ant-plant mutualisms are common interactions in which plants provide extra-floral nectar and/or ``housing'' to ants which in turn defend them from herbivores\cite{Bronstein1998, Bronstein2006}. 
Ants are highly ubiquitous\cite{Schultheiss2022} and vary significantly in abundances\cite{Schultheiss2022}, types of rewards offered\cite{Beattie1985}, and quality of interaction with their plant partners\cite{Rico-Gray1989,Palmer2010}, making these interactions great systems to study the effects of partner diversity within mutualisms. 
While these interactions have been well studied\cite{Ness2006,Beattie1985,Schultheiss2022} in the literature, few have considered how diversity within ant defender guilds\cite{Stanton2013} or temporal fluctuations in partner interactions\cite{} affect the overall benefits of mutualism for the plant partner.
